A full approximation storage multilevel method \\ for nonlinear variational inequalities

### 2023-06-22 paper/fascd.tex
results for classical obstacle problem example

\usepackage{times,bm,bbm,empheq,fancyvrb,graphicx,amsthm,amssymb}
\usepackage[dvipsnames]{xcolor}
\usepackage{longtable,booktabs,xspace}

@@

\usepackage{tikz}
\usetikzlibrary{decorations.pathreplacing}
\usetikzlibrary{graphs}

@@

Now suppose we repeatedly apply \pr{fascd-vcycle}, starting with $w^{J,0} \in \mathcal{K}^J$, generating (fine-level) iterates $w^{J,k} \in \mathcal{K}^J$.  Let $\id{atol}\ge 0$, $\id{rtol}\ge 0$ be given tolerances.  Then the criterion used in the results (Section \ref{sec:results}) is 
\begin{equation}
\|\rSS(w^{J,k})\|_2 < \id{atol} \qquad \text{or} \qquad \frac{\|\rSS(w^{J,k})\|_2}{\|\rSS(w^{J,0})\|_2} < \id{rtol}
\end{equation}
A step tolerance, based on $\|w^{J,k} - w^{J,k-1}\|_2$, could also be used, but we did not find it necessary.

@@

\section{Results and discussion} \label{sec:results}

The FASCD V-cycle and F-cycle (Algorithms \ref{alg:fascd}, \ref{alg:fascd-fcycle}) have been implemented in Python using the Firedrake library \cite{Rathgeberetal2016}.  In this Section we show numerical results corresponding to several VI examples from Section \ref{sec:vi}, with a focus on the scaling of FASCD iterations as the mesh resolution increases.

\begin{example} \label{ex:results:classical}
Consider the 2D classical, unilateral obstacle problem, namely Example \ref{ex:plaplacian} with $p=2$ and admissible set $\mathcal{K} = \{v \ge \psi\} \subset W^{1,2}(\Omega)$.  Suppose $\Omega$ is a square.  We consider two problems with very different coincidence (active) sets, the ``ball'' problem with $\psi$ equal to an upper hemisphere \cite[Chapter 12]{Bueler2021}, and the ``spiral'' problem in which $\psi$ is constructed to produce a nontrivial coincidence set \cite[Example 7.1.1]{GraeserKornhuber2009}.  FIXME show table of results
\end{example}

%REGENERATE using pefarrell-fascd/examples/
%$ timer python3 drape.py -prob ball|spiral -levs 9 -o ball|spiral.pvd -fascd_rtol 1.0e-6 -fascd_atol 1.0e-6 -fascd_cycle_type full
%$ paraview ball|spiral.pvd
%use preset "xray" and invert colors
\begin{figure}[ht]
\begin{center}
\includegraphics[width=0.35\textwidth]{fixfigs/ball-set.png} \qquad \includegraphics[width=0.35\textwidth]{fixfigs/spiral-set.png}
\end{center}
\caption{Coincidence sets $u=\psi$ for the ball and spiral problems.}
\label{fig:ballspiral}
\end{figure}

@@

% figures in the following visualization were generated using branch elbueler/simplefigs
% of git@bitbucket.org:pefarrell/fascd.git, and the following commands:
%   cd examples/
%   python3 plap1d.py -figures -fascd_max_it 2 -fascd_cycle_type multiplicative -mcoarse 2
\begin{figure}[ht]
\begin{center}
\input{tikz/imagesvcycle.tex}
\end{center}
\caption{Visualization of one iteration of \textsc{FASCD-VCYCLE} (Algorithm \ref{alg:fascd}) for Example FIXME, with $p=1.5$, $J=3$, and a coarsest mesh of two elements.  Solutions ($w^j$, $y^j$, $z^j$) are lines while obstacles ($\underline{\gamma}^j$, $\underline{\phi}^j$, $\underline{\chi}^j$) are strong dots.}
\label{fig:imagesvcycle}
\end{figure}

@@

@@

\section{From slack, to not lose} \label{app:slack}

FIXME remove this appendix, after using it in the introduction and discussion etc.

@@

So the difference between FASCD and \cite{GraeserKornhuber2009} are: (1) FAS coarse corrections, (2) no attempt to do monotone/truncated multigrid which would be a mess (I think) across the FE infrastructure, (3) clarity on what sets you can use in $V(\nu_1,\nu_2)$ cycles, with bigger up-smoother sets in particular, and most importantly (4) implementation within the mature Firedrake/PETSc library framework where one can easily hook up an active-set Newton step as a smoother.

>>>>

\usepackage{times,bm,bbm,empheq,verbatim,fancyvrb,graphicx,amsthm,amssymb}
\usepackage[dvipsnames]{xcolor}
\usepackage{booktabs,multirow,xspace}

@@

\usepackage{tikz}
\usetikzlibrary{decorations.pathreplacing}
\usetikzlibrary{graphs,quotes}

@@

Now suppose we repeatedly apply \pr{fascd-vcycle}, starting with $w^{J,0} \in \mathcal{K}^J$, generating (fine-level) iterates $w^{J,k} \in \mathcal{K}^J$.  Let $\id{atol}\ge 0$, $\id{rtol}\ge 0$ be given tolerances.  Then the criterion used in the results (Section \ref{sec:results}) is 
\begin{equation}
\|\rSS(w^{J,k})\|_2 < \id{atol} \qquad \text{or} \qquad \frac{\|\rSS(w^{J,k})\|_2}{\|\rSS(w^{J,0})\|_2} < \id{rtol} \label{eq:stoppingcriterion}
\end{equation}
A step tolerance, based on $\|w^{J,k} - w^{J,k-1}\|_2$, could also be used, but we did not find it necessary.

@@

\section{Results} \label{sec:results}

Algorithms \ref{alg:fascd} and \ref{alg:fascd-fcycle} were implemented in Python using the Firedrake library \cite{Rathgeberetal2016}.  In this Section we show numerical results corresponding to several VI examples from Section \ref{sec:vi}, with a focus on the scaling of FASCD iterations as the mesh resolution increases.

\begin{example} \label{ex:results:classical}
Consider the 2D classical, unilateral obstacle problem, namely Example \ref{ex:plaplacian} with $p=2$ and admissible set $\mathcal{K} = \{v \ge \psi\} \subset W^{1,2}(\Omega)$, with $\Omega$ square.  We consider two problems with very different coincidence (active) sets $\{x\in\Omega \,:\, u(x)=\psi(x)\}$.   The first is a standard ``ball'' problem, with an upper hemisphere obstacle $\psi$ \cite[Chapter 12]{Bueler2021}, and the second is the ``spiral'' problem in which $\psi$ is constructed to produce a more interesting coincidence set \cite[Example 7.1.1]{GraeserKornhuber2009}.  For this experiment we generated uniform triangular meshs, from a coarsest mesh of 32 triangles and $N=4^2$ nodes, to a finest mesh of $N=2048^2 \approx 4.2 \times 10^{6}$ nodes after nine standard refinements.  (The V- and F-cycles descended to the $4\times 4$ mesh.)  An RS$+$CG$+$ICC symmetric smoother (Section \ref{sec:implementation}) was applied, with fixed three iterations, and stopping tolerances \texttt{rtol} $=$ \texttt{atol} $= 10^{-6}$ in \eqref{eq:stoppingcriterion}.
% FIXME set atol *small* and re-run

Table \ref{tab:results:classical} shows the resulting iterations.  The RS Newton method, applied by itself as the solver with direct (LU) solution of the Newton steps, shows rapidly increasing iterations, approximately proportionally to the resolution, as expected; the highest resolution cases were not attempted with this non-scalable solver.  FASCD V-cycle iterations increase slowly, roughly logarithmically with the resolution.  For F-cycles we counted the number of V-cycles after the ramp (Section \ref{sec:implementation}), which shows constant iterations.  The spiral problem appears to be only slightly more challenging for these solvers.  Note that three V-cycle iterations in the F-cycle application, which suffices to solve either problem at any resolution, does work proportional to approximately FIXME smoother applications on the finest level, so these problems are solved with nearly textbook multigrid efficiency despite the low solution regularity.
\end{example}

%REGENERATE using pefarrell-fascd/examples/
%$ timer python3 drape.py -prob ball|spiral -levs 9 -o ball|spiral.pvd -fascd_rtol 1.0e-6 -fascd_atol 1.0e-6 -fascd_cycle_type full
%$ paraview ball|spiral.pvd
%use preset "xray" and invert colors
\begin{figure}[ht]
\begin{center}
\includegraphics[width=0.35\textwidth]{fixfigs/ball-set.png} \qquad \includegraphics[width=0.35\textwidth]{fixfigs/spiral-set.png}
\end{center}
\caption{Coincidence sets for the ball and spiral problems in Example \ref{ex:results:classical}.}
\label{fig:ballspiral}
\end{figure}

\newcommand{\XX}{{\sffamily X}}
% see pefarrell/examples/drape.py:
%   $ python3 drape.py -prob PROB -levs LEVS
% for PROB=ball,spiral and LEVS=1,...,10
% note mcoarse=4, rtol=atol=1e-6
% smoother is vinewtonrsls + 3 x (cg + icc)
\begin{table}[ht]
\begin{tabular}{lr rrr rrr}
\multirow{2}{*}{\emph{levels}} & \multirow{2}{*}{$N$} & \multicolumn{3}{c}{\,\emph{ball}} & \multicolumn{3}{c}{\,\emph{spiral}} \\ \cmidrule(lr){3-5} \cmidrule(lr){6-8}
             && \, V-cycle & F-cycle & RS only & \,V-cycle & F-cycle & RS only \\ \cmidrule{1-8}
 1 &    $4^2$ &  1 &  1 &  1  &  1 &  1 &  1 \\
 2 &    $8^2$ &  2 &  1 &  2  &  2 &  1 &  2 \\
 3 &   $16^2$ &  2 &  2 &  3  &  3 &  2 &  4 \\
 4 &   $32^2$ &  3 &  2 &  6  &  3 &  2 &  7 \\
 5 &   $64^2$ &  3 &  2 &  8  &  4 &  3 &  9 \\
 6 &  $128^2$ &  4 &  2 & 17  &  4 &  3 & 16 \\
 7 &  $256^2$ &  4 &  2 & 33  &  4 &  3 & 27 \\
 8 &  $512^2$ &  5 &  2 & 65  &  5 &  3 & 48 \\
 9 & $1024^2$ &  5 &  2 & \XX &  5 &  3 & \XX \\
10 & $2048^2$ &  5 &  2 & \XX &  5 &  3 & \XX
\end{tabular}

\bigskip
\caption{Iterations for FASCD V-cycle and F-cycle algorithms in Example \ref{ex:results:classical}.}
\label{tab:results:classical}
\end{table}

@@

% figures in the following visualization were generated using branch elbueler/simplefigs
% of git@bitbucket.org:pefarrell/fascd.git, and the following commands:
%   cd examples/
%   python3 plap1d.py -figures -fascd_max_it 2 -fascd_cycle_type multiplicative -mcoarse 2
\begin{figure}[ht]
\begin{center}
\input{tikz/imagesvcycle.tex}
\end{center}
\caption{Visualization of one FASCD V-cycle iteration for Example \ref{ex:results:plap}, with $p=1.5$, $J=3$, and a coarsest mesh of two elements.  Solutions (i.e.~$w^j$, $y^j$, $z^j$) are lines while obstacles ($\underline{\gamma}^j$, $\underline{\phi}^j$, $\underline{\chi}^j$) are strong dots.  The zero function (dotted) is the admissible initial iterate for $y^j$ and $z^j$.  Vertical scale varies.}
\label{fig:imagesvcycle}
\end{figure}

@@

\section{Discussion and Conclusion} \label{sec:discussion}

FIXME

@@

\begin{comment}
%From slack, to not lose

@@

So the difference between FASCD and \cite{GraeserKornhuber2009} are: (1) FAS coarse corrections, (2) no attempt to do monotone/truncated multigrid which would be a mess (I think) across the FE infrastructure, (3) clarity on what sets you can use in $V(\nu_1,\nu_2)$ cycles, with bigger up-smoother sets in particular, and most importantly (4) implementation within the mature Firedrake/PETSc library framework where one can easily hook up an active-set Newton step as a smoother.
\end{comment}
